The Concurrent Calculi Formalisation Benchmark

### 2023-10-15 challenges/coinduction.tex
Add proof of coinduction challenge.

For this challenge, we adopt a fragment of the untyped $\pi$-calculus
that includes process replication.
% The setting for this challenge is an untyped calculus of
% communicating systems with replication of processes, the syntax and
% semantics of which are presented below.
The objective of this challenge is to % prove % a context theorem for
draw a formal connection between strong barbed congruence and strong
barbed bisimilarity.  % I.e., The context lemma
The result establishes that two processes are strong barbed congruent
if and only if the processes obtained by applying a substitution to
them and composing them in parallel with an arbitrary process are
strongly barbed bisimilar.  Which, to reiterate, entails the use of
coinduction due to the presence of infinite behaviour.

@@

\subsubsection{Challenge}
The objective of this challenge is to prove a theorem that shows that
making strong barbed bisimilarity sensitive to substitution and
parallel composition is equivalent to strong barbed
congruence. Formally,
\begin{theorem}
  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
  substitution \( \sigma \),
  \(
  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
  \).
\end{theorem}

>>>>

For this challenge, we adopt a fragment of the untyped $\pi$-calculus
that includes process replication.
% The setting for this challenge is an untyped calculus of
% communicating systems with replication of processes, the syntax and
% semantics of which are presented below.
The objective of this challenge is to % prove % a context theorem for
draw a formal connection between strong barbed congruence and strong
barbed bisimilarity.  % I.e., The context lemma
The result establishes that two processes are strong barbed congruent
if the processes obtained by applying a finite number of substitutions to
them and composing them in parallel with an arbitrary process are
strongly barbed bisimilar.  Which, to reiterate, entails the use of
coinduction due to the presence of infinite behaviour.

@@

\subsubsection{Challenge}
The objective of this challenge is to prove a theorem that shows that
making strong barbed bisimilarity sensitive to substitution and
parallel composition is enough to show strong barbed
congruence.
To prove the theorem, we will use an \emph{up-to technique}.
A relation \( \mathcal{S} \) is called a \emph{strong barbed bisimulation up to \( \sbbisim{}{} \)} if, whenever \( (P,Q) \in \mathcal{S} \), the following conditions hold:
\begin{enumerate}
\item \( \observable{P}{\mu} \) if and only if \( \observable{Q}{\mu} \).
\item if \( \transition{P}{\Atau}{P'} \) then \( \transition{Q}{\Atau}{Q'} \) for some \( Q' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
\item if \( \transition{Q}{\Atau}{Q'} \) then \( \transition{P}{\Atau}{P'} \) for some \( P' \) with \( \uptosbb{P'}{\mathcal{S}}{Q'} \).
\end{enumerate}
These relations are useful owing to the following lemma:
\begin{lemma}\label{lemma:up-to}
  If \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \), \( (P,Q) \in \mathcal{S} \) implies \( \sbbisim{P}{Q} \).
\end{lemma}
\begin{proof}
  We check that \( \uptosbb{}{\mathcal{S}}{} \) is a strong barbed bisimulation and is thus included in \( \sbbisim{}{} \).
\end{proof}

We are now ready to prove the theorem:
\begin{theorem}
  \( \sbcong{P}{Q} \) if, for any process \( R \) and
  substitution \( \sigma \),
  \(
  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
  \).
\end{theorem}
\begin{proof}
  Since \( \sbcong{}{} \) is the largest congruence included in \( \sbbisim{}{} \), it suffices to show that if \( \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}} \) for any \( R \) and \( \sigma \), then \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}} \) for any \( C \), \( R \) and \( \sigma \).
  We proceed by induction on \( C \).
  \begin{description}
  \item[\( C = \Pin{x}{z}{C'} \)] Let \( \mathcal{S} = \{ (\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}, \Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}) \mid \text{$R$ and $\sigma$ arbitrary} \}~\cup \sbbisim{}{} \).
    We can easily check that \( \mathcal{S} \) is a strong barbed bisimulation, noting that \( \sbbisim{}{} \) is preserved by restriction and is contained in \( \mathcal{S} \).
  \item[\( C = \Ppar{C'}{S} \)] Then by the induction hypothesis,
    \begin{equation*}
      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}}{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{(\Ppar{\applysubst{\sigma}{S}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
    \end{equation*}
    for any \( R \) and \( \sigma \).
  \item[\( C = \Pres{z}{C'} \)] Then by the induction hypothesis we have \( \sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R}}{\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R}} \) for any \( R \) and \( \sigma \).
    Without loss of generality, we assume that \( z \notin \freenames{R} \cup \names{\sigma} \).
    Then, using that \( \sbbisim{}{} \) is preserved by restriction, we have
    \begin{equation*}
      \sbbisim{\sbbisim{\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}}{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{P}}}{R})}}}{\sbbisim{\Pres{z}{(\Ppar{\applysubst{\sigma}{\applyctx{C'}{Q}}}{R})}}{\Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}}}
    \end{equation*}
  \item[\( C = \Preplicate{C'} \)] Let \( \mathcal{S} = \{ (\Ppar{\applysubst{\sigma}{\applyctx{C}{P}}}{R}, \Ppar{\applysubst{\sigma}{\applyctx{C}{Q}}}{R}) \mid \text{$R$ and $\sigma$ arbitrary} \} \).
    Using \cref{lemma:up-to}, it suffices to show that \( \mathcal{S} \) is a strong barbed bisimulation up to \( \sbbisim{}{} \).
    To this end, let
    \begin{align*}
      A &= \applysubst{\sigma}{\applyctx{C'}{P}}, & A' &= \applysubst{\sigma}{\applyctx{C}{P}} \\
      B &= \applysubst{\sigma}{\applyctx{C'}{Q}}, & B' &= \applysubst{\sigma}{\applyctx{C}{Q}} ,
    \end{align*}
    noting that \( A' = \Preplicate{A} \) and \( B' = \Preplicate{B} \).

    Suppose \( \transition{\Ppar{R}{A'}}{\Atau}{S} \) for some \( S \).
    Then we can show by a case analysis on the derivation of this transition that there exists a \( T \) such that \( \transition{\Ppar{R}{(\Ppar{A}{A})}}{\Atau}{T} \) and \( \sbbisim{S}{\Ppar{T}{A'}} \).
    Using the induction hypothesis twice, we note that \( \sbbisim{\Ppar{R}{B'}}{\Ppar{R}{\Ppar{(\Ppar{A}{A})}{B'}}} \).
    Since by rule \TirName{Par-L}, \( \transition{\Ppar{\Ppar{R}{(\Ppar{A}{A})}}{B'}}{\Atau}{\Ppar{T}{B'}} \), there must thus exist a \( U \) such that \( \transition{\Ppar{R}{B'}}{\Atau}{U} \), \( \sbbisim{U}{\Ppar{T}{B'}} \) and \( \uptosbb{S}{\mathcal{S}}{U} \) as required.
    The proof for \( \transition{\Ppar{R}{B'}}{\Atau}{S} \) is analogous.
  \end{description}
  The remaining cases are similar.
\end{proof}

### 2023-10-15 paper/paper.tex
Add proof of coinduction challenge.

For this challenge, we adopt a fragment of the untyped $\pi$-calculus
that includes process replication.
% The setting for this challenge is an untyped calculus of
% communicating systems with replication of processes, the syntax and
% semantics of which are presented below.
The objective of this challenge is to % prove % a context theorem for
draw a formal connection between strong barbed congruence and strong
barbed bisimilarity.  % I.e., The context lemma
The result establishes that two processes are strong barbed congruent
if and only if the processes obtained by applying a substitution to
them and composing them in parallel with an arbitrary process are
strongly barbed bisimilar.
The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.

@@

\subsubsection{Challenge.}
The objective of this challenge is to prove a theorem that shows that
making strong barbed bisimilarity sensitive to substitution and
parallel composition is equivalent to strong barbed
congruence. Formally,
\begin{theorem}
  \( \sbcong{P}{Q} \) if and only if for any process \( R \) and
  substitution \( \sigma \),
  \(
  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
  \).
\end{theorem}

@@

\section{Going beyond the challenge problems.}\label{sec:going-beyond}

>>>>

For this challenge, we adopt a fragment of the untyped $\pi$-calculus
that includes process replication.
% The setting for this challenge is an untyped calculus of
% communicating systems with replication of processes, the syntax and
% semantics of which are presented below.
The objective of this challenge is to % prove % a context theorem for
draw a formal connection between strong barbed congruence and strong
barbed bisimilarity.  % I.e., The context lemma
The result establishes that two processes are strong barbed congruent
if the processes obtained by applying a finite number of substitutions to
them and composing them in parallel with an arbitrary process are
strongly barbed bisimilar.
The key issue of this challenge is the coinductive reasoning about the infinite behaviours of the replication operator.

@@

\subsubsection{Challenge.}
The objective of this challenge is to prove a theorem that shows that
making strong barbed bisimilarity sensitive to substitution and
parallel composition is enough to show strong barbed
congruence. Formally,
\begin{theorem}
  \( \sbcong{P}{Q} \) if for any process \( R \) and
  substitution \( \sigma \),
  \(
  \sbbisim{\Ppar{\applysubst{\sigma}{P}}{R}}{\Ppar{\applysubst{\sigma}{Q}}{R}}
  \).
\end{theorem}

@@

\section{Going beyond the challenge problems}\label{sec:going-beyond}
